The missing puzzle piece: Contextual insights for enhanced pharmaceutical supply chain forecasting

### 2024-11-02 manuscript/04-results.qmd
first draft ready

```{r}
#| include: false
#| label: restults
library(readr)
fc_accuracy_overall__without_predictor <- read_csv("../result/fc_accuracy_overall__without_predictor.csv")
fc_accuracy_predictor_overall <- read_csv("../result/fc_accuracy_predictor_overall.csv")
fc_accuracy_overall <- read_csv("../result/fc_accuracy_all_item.csv")
data_visualisation <- read.csv("../rscript/data_visualisation.R")
```

from the total of 37 key pharmaceuticals selected and tested with the  different forecasting models arima showed that a better forecast accuracy for most of the pharmaceuticals followed by mean and  regresstion models with RMSSE value of 0.850, 0.942 and 0.945 respectively (Table 1).

```{r}
library(kableExtra)
library(dplyr)  # Load the dplyr package
fc_accuracy_cleaned <- na.omit(fc_accuracy_overall__without_predictor)
fc_accuracy_table <- fc_accuracy_cleaned %>%
  dplyr::select(.model, RMSE, MAE, MASE, RMSSE, winkler, CRPS)  # Explicitly specify dplyr::select
kbl(fc_accuracy_table, booktabs = TRUE, escape = FALSE) %>%
  kable_styling(latex_options = c("hold_position"), full_width = FALSE)
```

Then we tested the models by adding different predictors which affects the sales of the pharmaceuticals like physical count, budget release and budget closure by the government, COVID 19, campaign in to the models as predictors which will affects the sales of the key pharmaceuticals products and the result showed that still arima have a better forecast accuracy with RMSSE of 0.850 (Table 2).

```{r}
#| label: result3
#| tbl-cap: "forecast accuarcy of selected pharmaceuticals in EPSS 2018-2022 after considering predictors"
fc_accuracy_predictor_overall %>% 
  kbl(booktabs = TRUE, escape = FALSE) |>
  kable_styling(latex_options = c("hold_position"),full_width = F)
```
```{r}
#| label: result2
#| fig-cap: "forecast accuarcy of selected pharmaceuticals in EPSS 2018-2022 after considering predictors"
library(ggplot2)  # Load the ggplot2 package
library(forcats)  # Load the forcats package
fc_accuracy_overall %>%
  select(`.model`, item, RMSSE) %>% 
  ggplot(mapping = aes(x = RMSSE, y = fct_reorder(`.model`, RMSSE, .fun = median))) +
  geom_boxplot()
```

Individual item with different forecasting models 
We tested 10 models namely arima, arima with predictor, combination, exponential smoothing, mean, naive, regression, regression with predictor, snaive and stlf. Individually on the selected key pharmaceuticals and the result indicated that 11 (32.43%) of the selected pharmaceuticals have better forecast accuracy for arima predictor models, followed by naive 6 (16.21%), mean, regresstion, arima, and STLF methods each of them with 3 (8.10%) (Fig 2)

```{r}
# Sample data (replace with your actual data)
pharmaceuticals <- c("ARIMA Predictor", "Naive", "Mean", "Regression", "ARIMA", "STLF")
accuracy_percentages <- c(32.43, 16.21, 8.10, 8.10, 8.10, 8.10)

# Create a data frame
data <- data.frame(Pharmaceuticals = pharmaceuticals, Accuracy_Percentage = accuracy_percentages)
# Create a bar chart
library(ggplot2)

ggplot(data, aes(x = reorder(Pharmaceuticals, -Accuracy_Percentage), y = Accuracy_Percentage)) +
  geom_bar(stat = "identity", fill = "blue") +
  labs(x = "Pharmaceutical Forecasting Models", y = "Accuracy Percentage") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

@@

Those pharmaceuticals who had a better forecast accuracy with Arima predictors are Amoxacillin 500mg capsule, Anti RhD, Artehmether + Lumefantrine, Atenolol, Atrovastatin 200mg, Ferrosu sulphate with follic acid, frusemide injection, Insuline Iosphane bipahsic, Lamivudine +Efaverienze + Tenofovir, Lidocaine + HCL and Propyl thiro uracil.The  six pharmaceuticasl which showed a better forecast accuracy with naive models were Ciprofloxacillin 500mg tablet, Magnesoum sulphate injection, Omeprazole capsule, Omeprazole injection, Rapid diagnostic test, Sulphamethaxone + Trimetoprine tablet.

Discussion:

The objective of this study was to evaluate the forecasting accuracy of different 
models for key selected pharmaceutical products from the Ethiopian Pharmaceutical 
Supply Services (EPSS). The analysis encompassed a total of 37 key pharmaceutical 
products, and the results highlight the performance of various forecasting models in 
predicting sales for these products. In the initial assessment of forecasting models, 
the ARIMA model emerged as the most accurate for the majority of the pharmaceutical 
products, exhibiting a Root Mean Squared Scaled Error (RMSSE) value of 0.850. This 
finding suggests that ARIMA has a strong predictive capability and outperforms other 
models, including the mean and regression models. The superiority of the ARIMA models 
over the Moving Average might be explained by  most of the data set have seasonal 
patern and MA  is one of the simplest prediction techniques for making projections 
about time-series without a noticeable seasonal pattern [@chopra2001supply].The 
superiority of the ARIMA model underscores its suitability for pharmaceutical sales 
forecasting within the context of EPSS. Similar findings were reported in forecasting 
ethanol demand in India where ARIMA model outperforms other models like linear and 
non-linear regression models [@dey2023forecasting].

To enhance the predictive power of the models, additional predictors were incorporated 
into the analysis. Predictors such as physical count, budget release and closure by the government, the impact of COVID-19, and promotional campaigns were considered as 
potential drivers of pharmaceutical sales. Despite the inclusion of these predictors, 
the ARIMA model maintained its superior forecasting accuracy with an RMSSE of 0.850. 
This resilience further validates the robustness of the ARIMA model and its ability to 
capture complex interactions between variables influencing pharmaceutical sales.

To delve deeper into the individual performance of the forecasting models, a 
comprehensive evaluation was conducted on each of the 37 pharmaceutical products. Among the ten models tested, ARIMA with predictors exhibited the highest forecast accuracy 
for 32.43% of the products. This was followed by the naive model, which performed well 
for 16.21% of the products. Notably, a few pharmaceuticals demonstrated a strong fit 
with specific models. For instance, Amoxicillin 500mg capsule, Anti RhD, Artehmether + 
Lumefantrine, and others exhibited superior forecast accuracy when utilizing the ARIMA 
model with predictors.

Similarly, other pharmaceuticals displayed better forecasting outcomes with alternative models. Notably, Ciprofloxacin 
500mg tablet, Magnesium sulfate injection, Omeprazole capsule, and others showed 
improved accuracy with the naive model. These findings underscore the importance of 
tailoring the choice of forecasting model based on the characteristics and 
dynamics of individual pharmaceutical products.This finding confirmed that There is not
a single best technique to solve 
time-series forecasting problems [@zhang2007quarterly]. In order to deal with 
time-series forecasting, each problem might be solved with a different approach [@ensafi2022time].

In conclusion, this study highlights the pivotal role of accurate forecasting in the 
pharmaceutical supply chain management. The ARIMA model, particularly when integrated 
with relevant predictors, emerges as a powerful tool for 
predicting pharmaceutical sales within the EPSS. However, the selection of the most 
appropriate model should consider the specific attributes of each pharmaceutical 
product. Further research could explore the optimization of model parameters and the 
incorporation of additional contextual variables to enhance forecasting accuracy and 
supply chain efficiency.

Limitation

One notable limitation of this study revolves around the availability and quality of 
the data used for analysis. While we employed a comprehensive dataset from the EPSS, it is essential to acknowledge that the accuracy of forecasting models is heavily 
contingent on the quality, granularity, and completeness of historical sales data. 
Variations in data recording practices, potential errors, missing values, or 
inconsistencies within the dataset could impact the precision of the forecasts 
generated by the employed models. Moreover, the inherent complexity of pharmaceutical 
demand, influenced by a myriad of external factors such as socioeconomic changes,stock 
outs, prolonged procurement lead time, rationing of avaiable products, pushing of 
pharmaceuticals to different health facility without demand, healthcare policies, and 
unforeseen events, may introduce an additional layer of uncertainty. It is important to recognize that despite our best efforts to address data-related challenges, the 
robustness and reliability of our forecasting outcomes may still be subject to the 
limitations inherent in the original data.

>>>>

```{r}
#| include: false
#| label: restults
library(readr)
point_accuracy_h <- read_rds("../result/accuracy_h.rds") |> janitor::clean_names() |> select(item, h,model,mase,rmsse)
prob_accuracy_h <- read_rds("../result/accuracy_h.rds") |> janitor::clean_names() |> select(item, h,model,crps_skill,winkler_skill=skill,crps,winkler)

point_accuracy_id <- read_rds("../result/accuracy_id.rds")|> janitor::clean_names() |> select(item, id,model,mase,rmsse)
prob_accuracy_id <- read_rds("../result/accuracy_id.rds")|> janitor::clean_names() |> select(item, id,model,crps_skill,winkler_skill=skill,crps,winkler)

point_accuracy_horizon <- read_rds("../result/point_accuracy_horizon.rds") |> select(item, h=horizon,model,mase,rmsse)
prob_accuracy_horizon <- read_rds("../result/prob_accuracy_horizon.rds") |>  janitor::clean_names() |> select(item, h=horizon,model,crps_skill=crps_skill_score,winkler_skill=winkler_skill_score,crps,winkler=winkler_score)

point_accuracy_origin <- read_rds("../result/point_accuracy_origin.rds") |> janitor::clean_names() |> select(item, id=origin,model,mase,rmsse)
prob_accuracy_origin <- read_rds("../result/prob_accuracy_origin.rds") |> janitor::clean_names() |> select(item, id=origin,model,crps_skill=crps_skill_score,winkler_skill=winkler_skill_score,crps,winkler=winkler_score)

point_accuracy_h_all <- point_accuracy_h |> bind_rows(point_accuracy_horizon |> mutate(h= as.integer(h)))
prob_accuracy_h_all <- prob_accuracy_h |> bind_rows(prob_accuracy_horizon |> mutate(h= as.integer(h)))

point_accuracy_origin_all <- point_accuracy_id |> bind_rows(point_accuracy_origin |> mutate(id= as.integer(id)))
prob_accuracy_origin_all <- prob_accuracy_id |> bind_rows(prob_accuracy_origin |> mutate(id= as.integer(id)))

@@

In this section, we compare the forecasting performance of various approaches, examining models that incorporate expert-identified business context predictors versus those that rely solely on historical consumption data. Point forecast performance is reported using MASE and MSSE, while probabilistic forecast accuracy is reported using CRPS.

The forecasting performance is reported in @fig-point, in which the average forecast accuracy over forecast horizon and across all products is calculated for each origin. we also report the ditribution of accurcy metrics  across all rollog origins, average across forecast horizon and all products. this shows how mdoels varies in roviding accuracy across different origins. further rserach would be required to understand how error analysis across forecast hrizon and across differne origins could be used as afiagnos!
y-axis shows the models sorted based on ther values of mase and rmsse, when the one i the bottom has lowest error. this modesis TimeGpt, which is the foundational time sei models.

@fig-point also indicate that predictors collected through interacting with experts with domaine knoeldge improves point forecast accuracy for all models, highlighting the essential needs to systematically colelct such information alognside the tarsactional data about consuption.

We observed that the LSTM model had the highest error compared to all other models. While LSTMs are powerful for capturing complex, non-linear temporal patterns, they may underperform in certain scenarios, especially when compared to simpler time series forecasting models. This underperformance can be attributed to overfitting, as LSTMs have a large number of parameters and require substantial amounts of data to generalize effectively. In contrast, simpler models like ARIMA with exogenous variables often handle straightforward trends or seasonality more effectively and may outperform LSTMs when long-term dependencies or non-linearity in the data are not significant.

```{r}
#| label: fig-point
#| fig-width: 7
#| fig-height: 5
#| fig-cap: "Distribution of point forecast accuracy across different origins, averaged across the forecat horizon and all products. The total number of months used to calculate the accuracy in the test set is 12 months for each product. MASE and MSSE are relative to the corresponding values for the training set."
#| fig-pos: "H"

p_mase <- point_accuracy_origin_all |> ggplot(aes(x= mase, y =  fct_reorder(model, mase)))+
  geom_boxplot()+
  labs(y = "Models")+
  ggthemes::theme_few()

p_rmsse <- point_accuracy_origin_all |> ggplot(aes(x= rmsse, y =  fct_reorder(model, mase)))+
  geom_boxplot()+
  labs(y = "Models")+
  ggthemes::theme_few()

point_patch <- p_mase/p_rmsse

point_patch+ggthemes::theme_few()

```{r}
#| label: fig-prob
#| fig-width: 9
#| fig-height: 3
#| fig-cap: "Distribution of probabilistic forecast accuracy across different origins, averaged across the forecat horizon and all products. The total number of months used to calculate the accuracy in the test set is 12 months for each product."
#| fig-pos: "H"

prob_accuracy_origin_all |> ggplot(aes(x= crps, y =  fct_reorder(model, crps)))+
  geom_boxplot()+
  labs(y = "Models")+
  ggthemes::theme_few()

```{r}
#| label: fig-horizon
#| fig-width: 11
#| fig-height: 5
#| fig-cap: "Average accuracy by month for 6 month. The total number of months used to calculate the accuracy in the test set is 12 for each product."
#| fig-pos: "H"

p1h <- point_accuracy_h_all |> group_by(model,h) |> summarise(mase=mean(mase) , .groups = "drop")|> filter(model != "lstm", model != "lstm_reg") |>  ggplot(aes(x = h, y = mase, colour =model))+
  geom_point()+
  geom_line()+
  ggthemes::theme_few()+
  ggthemes::scale_color_colorblind()+
  labs(x= " Forecast horizon")+
  theme(legend.position="none")

p2h <- point_accuracy_h_all |> group_by(model,h) |> summarise(rmsse=mean(rmsse),.groups = "drop") |> filter(model != "lstm", model != "lstm_reg") |> 
  ggplot(aes(x = h, y = rmsse, colour =model))+
  geom_point()+
  geom_line()+
  ggthemes::theme_few()+
  ggthemes::scale_color_colorblind()+
  labs(x= " Forecast horizon")+
  theme(legend.position="bottom")

p3h <- prob_accuracy_h_all |> group_by(model,h) |> summarise(crps=mean(crps),.groups = "drop") |> filter(model != "lstm", model != "lstm_reg") |> 
  ggplot(aes(x = h, y = crps, colour =model))+
  geom_point()+
  geom_line()+
  ggthemes::theme_few()+
  ggthemes::scale_color_colorblind()+
  labs(x= " Forecast horizon")+
  theme(legend.position="none")

prob_patch <- (p1h|p2h|p3h)
prob_patch
```

@fig-prob illustrates the forecast distribution accuracy measured by the Continuous Ranked Probability Score (CRPS), which evaluates both forecasting calibration and interval sharpness. A smaller CRPS value indicates better overall performance. We observed that incorporating domain knowledge improved forecast accuracy for most models, enhancing not only point forecasts but also probabilistic forecasts. Notably, TimeGpt achieved the highest probabilistic forecast accuracy, while LSTM models had the lowest. This aligns with the findings related to point forecast accuracy, reinforcing the earlier explanations for these results.

In addition to reporting the forecast accuracy presented across different rolling origins, we also report the point and probabilistic forecast accuracy measures for each forecast horizon in @fig-horizon. this is the accuracy reorted as the average across rolling origins and all products fro each given horizon. For both the point forecast and distributional accuracy  we can see that models incoporating information collected by domain experts improves compared to counterparts that that uses only the historical comsuptions.

Overall, our results indicate that forecast produces using models with infromation colelcted via the domain expert engagement provides reliable forecasts and improves upon the models relying solely on historcial consuption.

Our analysis highlights the value of using foundational time series forecasting models like TimeGpt, particularly in developing countries and the global health and humanitarian sectors. In these contexts, the lack of advanced analytical skills and systematic data collection can hinder effective forecasting and decision-making.

The findings also emphasize the importance of incorporating relevant domain knowledge into forecasting models. In pharmaceutical supply logistics administration systems, data often consists solely of transactional records on consumption and distribution. Including exogenous variables such as administrative procedures, seasonal patterns, or conflicts, or any other relevant factor that could be identified by those with domain knowledge proved essential for reducing forecast errors. This approach improved both point and probabilistic forecast accuracy, enabling a more confident assessment of uncertainty. Therefore, the systematic collection of information about significant events and their impact on consumption is vital. Recording details of events such as policy changes, administration procedures, conflicts, and incorporating this information into forecasting models creates a comprehensive understanding of consumption. This practice enhances modeling reliability and allows institutions in developing countries and humanitarian organizations to better forecast demand, allocate resources effectively, and respond proactively. Establishing robust data collection systems is thus a critical step for strengthening forecasting capabilities and operational resilience.

## An illustration of probabilistic forecast for Pharmacuitical product consumption

In this section, we present an illustrative example of a probabilistic forecast for future consumption of _Sodium Chloride (Normal Saline)_ product. Due to the complexity of including such visualizations for all products, only one example is shown here. However, it is feasible to generate these plots for all products if needed.

In practice, point forecasts are commonly used despite their limitations, but they do not account for the inherent uncertainty associated with forecasts. The future is inherently uncertain, and effective planning requires considering alternative scenarios. Probabilistic forecasts offer a comprehensive approach by assigning likelihoods to a range of possible outcomes, recognizing that different consumption levels may occur with varying probabilities. The primary purpose of probabilistic forecasting, as illustrated in @fig-forecast-density-hstep, is to quantify and communicate uncertainty. This figure displays the forecast distribution of consumption over a 6-month horizon using a density plot. For each month within the forecast period, a separate distribution is generated. The plot also includes the point forecast alongside 80% and 90% prediction intervals to illustrate potential variability.

It is important to note that while point forecasts and prediction intervals can be derived from probabilistic forecasts, the reverse is not true. A single-point forecast cannot inherently provide the probabilistic context needed to capture the range of possible outcomes. Although prediction intervals can indicate a range of potential values, they do not convey the detailed probabilities of low or high consumption. This distinction highlights the value of probabilistic forecasting in supporting informed decision-making by offering a clearer view of future uncertainties.

```{r}
#| label: fig-forecast-density-hstep
#| fig-width: 9
#| fig-height: 4
#| fig-cap: "A graphical illustration of the forecast distribution of a pharmacutical product (i.e. total incidence attended) for the SB health board for a horizon of six month. For each month, we display the point forecast (black point), the histogram, and 80% (thick line) and 90% (thin line) prediction intervals. It also shows a portion of a historical time series as well as its fitted values."
#| fig-pos: "H"

amoxicillin <- med_qty |> filter(item == "Sodium Chloride (Normal Saline)") |> select(month,quantity) |> 
  as_tsibble(index = month)

fit <- amoxicillin |> filter_index(. ~ "2022 Jan") |>
  model(ets = ARIMA(quantity))

fcst <- fit |> forecast(h = 6)
fitted_ets <- fit |> augment()

ggplot(data = fcst, mapping = aes(x = month, ydist = quantity))+
  ggdist::stat_halfeye(alpha = .4)+
  geom_line(aes(y=.mean, colour ="Point Forecast"))+
  geom_line(aes(y = .fitted, colour ="Fitted"), data = filter_index(fitted_ets, "2021 Jan" ~ .))+
  geom_point(aes(y = .fitted, colour ="Fitted"), data = filter_index(fitted_ets, "2021 Jan" ~ .))+
  geom_line(aes(y = quantity, colour ="Actual"),data = filter_index(amoxicillin, "2021 Jan" ~ .))+
  geom_point(aes(y = quantity, colour ="Actual"),data = filter_index(amoxicillin, "2021 Jan" ~ .))+
  scale_color_manual(name=NULL,
                     breaks=c('Fitted', 'Actual',"Point Forecast"),
                     values=c('Fitted'='#E69F00', 'Actual'='#0072B2',"Point Forecast"="#000000"))+
  ggthemes::theme_few()
```

### 2024-11-02 manuscript/05-conclusion.qmd
first draft ready

# Conclusion {#sec-conclusion}

In this study, we embarked on a comprehensive analysis of pharmaceutical demand forecasting within the Ethiopian 
Pharmaceutical Supply Services (EPSS) context. By leveraging five  years of sales data for 37 key pharmaceutical 
products, we aimed to  enhance the accuracy of demand predictions and consequently  contribute to more effective supply 
chain management. Our findings  underscore the significance of accurate demand forecasting in  optimizing inventory 
levels, resource allocation, and overall supply chain operations.

The application of diverse forecasting models yielded valuable insights into their respective performance. Notably, the 
Autoregressive Integrated Moving Average (ARIMA) model emerged as a  robust choice, consistently exhibiting superior forecast accuracy  across a majority of the selected pharmaceutical products.  Additionally, the incorporation of predictor variables, such as  physical count, budget release and closure by the government,  COVID-19 impact, and campaign events, further underscored the  efficacy of the ARIMA model in accommodating external influences and delivering reliable forecasts.

The individualized assessment of forecasting models revealed specific strengths across pharmaceutical categories. ARIMA 
models with predictors demonstrated exceptional performance for pharmaceuticals including Amoxacillin 500mg capsule, Anti RhD, Artehmether + Lumefantrine, Atenolol, and others, affirming their utility in capturing complex demand patterns. 
Moreover, the study shed light on pharmaceuticals best suited for alternative models like naive, mean, and regression 
methods, contributing to a comprehensive understanding of forecasting applicability.

While these findings contribute significantly to the discourse on pharmaceutical demand forecasting, it's essential to 
acknowledge limitations. The accuracy of our forecasts is intrinsically linked to the quality and completeness of 
historical sales data, and the intricate interplay of external variables adds complexity to the forecasting landscape.

In conclusion, this study serves as a pivotal step towards enhancing pharmaceutical supply chain management through 
refined demand forecasting. The demonstrated success of ARIMA models, particularly when augmented with predictor 
variables, holds promise for fostering resilient and adaptable supply chains. As the pharmaceutical landscape continues 
to evolve, our insights encourage further exploration of advanced forecasting techniques and continuous refinement of 
strategies to bolster supply chain efficiency and ensure consistent availability of essential pharmaceuticals.

In light of the study's findings, several actionable recommendations emerge. First and foremost, the integration of the 
Autoregressive Integrated Moving Average (ARIMA) model with predictor variables stands out as a viable strategy for 
enhancing pharmaceutical demand forecasting accuracy within Ethiopian Pharmaceutical Supply Services (EPSS). This entails 
harnessing the predictive power of external factors to refine forecasting precision. To further advance forecasting 
practices, the exploration of additional predictor variables such as public health campaigns, disease outbreaks, and 
economic indicators could offer valuable insights into demand dynamics. Moreover, refining the granularity of historical 
sales data, possibly through finer temporal intervals and geographic diversity, holds potential for more nuanced and 
accurate predictions. Consideration could also be given to hybrid forecasting models that amalgamate ARIMA's short-term 
accuracy with machine learning's predictive prowess for long-term trends. Regular model evaluation, sensitivity analysis, and collaborative stakeholder engagement are recommended to ensure ongoing accuracy and alignment with changing demand patterns.

Capacitating EPSS staff through training programs in forecasting methodologies and data interpretation can empower the 
workforce to make informed decisions. Replication of this study in diverse healthcare settings and comparative analyses 
can validate the findings' applicability and adaptability. By embracing these recommendations, EPSS can fortify its 
pharmaceutical supply chain with resilient, efficient, and responsive operations, ultimately ensuring the consistent 
availability of vital pharmaceuticals.
  a

>>>>

# Conclusion and future reserach {#sec-conclusion}

In this study, we conducted an extensive analysis of pharmaceutical demand forecasting within the Ethiopian Pharmaceutical Supply Service (EPSS). Using five years of consumption data for 33 key pharmaceutical products, along with additional information gathered through collaboration with domain experts at EPSS, our goal is to contribute to enhancing the demand forecasting process and emphasize the importance of integrating domain knowledge into model building. This step is critical, as the phase of understanding the data and incorporating valuable contextual information is often overlooked. Too frequently, forecasting efforts rely solely on consumption data, leading to building models with little relevance to reality. Our approach highlights the necessity of including domain insights to construct more accurate and effective forecasting models.

Our findings highlight the importance of collecting and incorporating domain knowledge when building forecasting models. We evaluated both point and probabilistic forecasts using a range of models, from simple univariate approaches like ARIMA to complex models such as LSTM. Notably, we demonstrated that employing complex models does not necessarily lead to more accurate forecasts in the pharmaceutical supply context. Additionally, our research revealed that newly developed foundational time series forecasting models have potential applicability, particularly in environments where advanced analytical skills and knowledge are limited, as often seen in pharmaceutical supply chains in developing countries.

To advance forecasting practices further, exploring additional predictor variables—such as public health campaigns, disease outbreaks, and economic indicators—could provide valuable insights into demand dynamics. Improving the granularity of historical consumption data, through finer temporal intervals and geographic differentiation, also holds potential for more precise predictions. Additionally, hybrid forecasting models that combine judgmental point forecasts with probabilistic machine learning models could enhance predictive performance. Empowering EPSS staff through training in foundational time series models, data interpretation, and collection will enable informed decision-making and strengthen workforce capabilities. Replicating this study across diverse healthcare settings and conducting comparative analyses would help validate the adaptability and applicability of the findings. Additionally, it is recommended that pharmaceutical supply services systematically collect and maintain detailed records of events that may influence consumption, alongside consumption data. This practice is essential for refining demand models and enhancing forecast accuracy.

## Practical challenges and limitations

Despite dedicated efforts to engage with domain experts to better understand the data—and spending significant time over two weeks collaborating with them—we found that the process of interpreting data through domain knowledge is complex and presents unique challenges. These challenges include defining what constitutes domain knowledge and determining how best to incorporate it to enhance model reliability and accuracy. Below, we summarize some of these key challenges. One issue involved capturing accurate information from bin cards and expert review, particularly during periods of disruption like COVID-19 and conflicts. For instance, during the COVID-19 pandemic, experts noted a decrease in demand due to travel restrictions. However, despite the reduced demand, there were still significant distributions of pharmaceuticals from the EPSS central to various hubs and health facilities. This was because, in response to anticipated shortages, a political decision was made to push products to these facilities before the travel restrictions took full effect. The rationale was that if the facilities remained closed, patients would have no access to medications, necessitating the preemptive stockpiling of essential pharmaceuticals. This scenario illustrates how policy decisions can significantly impact supply chain data, making it challenging to model distribution patterns accurately. Similarly, experts indicated that conflicts would disrupt the distribution of pharmaceuticals. While conflict zones did indeed hinder transportation, distribution still occurred whenever roads were temporarily opened, even amidst on-going conflict. This created inconsistencies in the data, as periods of halted distribution were followed by rapid replenishment once access was restored. Such fluctuations add complexity to modelling the supply chain from the central EPSS to hubs and health facilities. Other challenges involved seasonal or periodic activities, such as the fiscal year-end stock counting. During this time, EPSS temporarily halts transactions to conduct a full inventory count, which can take anywhere from one to two months, depending on the circumstances. This inconsistency in the duration of inventory counts adds a layer of unpredictability to supply chain modelling.

## Future reserach

Following the current research, several promising areas could further advance knowledge and practical applications in this area:

- Future research could focus on developing intelligent systems that incorporate domain-specific knowledge into model building. This would involve creating algorithms capable of identifying relevant domain insights and integrating them effectively into model training processes. Such tools would bridge the gap between purely data-driven approaches and expert-knowledge-enhanced modeling, leading to more robust and contextually informed forecasts.

- The use of distribution-based representation of error metrics across different rolling origins or forecast horizons, such as boxplots, to showcase forecast accuracy provides essential insights into variability in accuracy at different point of time in data. Building on this, future work could aim at designing methodologies that identify diagnostic patterns within the distribution of errors. These approaches could suggest model improvements by analyzing error distributions and help practitioners to refine models based on error behavior.

- While current study focus on monthly aggregate consumption at the organizational level, there is potential in exploring more granular forecasts. Future research could investigate daily or weekly forecasting and examine the impacts at hierarchical levels, such as specific sites or health facilities. Investigating how fine-grained temporal and hierarchical forecasts influence inventory management and operational efficiency could yield valuable insights for reducing stockouts, minimizing costs and reducing waste.
